Batch, Off-Policy Apprencticeship Learning

### 2011-03-28 main.tex
Background de Matthieu

\typeout{Batch, Off-Policy Apprenticeship Learning}
\documentclass{article}
\usepackage{ijcai11}
\usepackage{times}
\usepackage{color}
% the following package is optional:
%\usepackage{latexsym} 
\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{subfigure}
%\usepackage{algorithm2e}
\usepackage{algorithm}
\usepackage{algorithmic}

@@

\section{Background}
\label{sec:back}
A \emph{Markov Decision Process} is a tuple $\{S,A,\gamma,R\}$ with :
\begin{itemize}
\item $S$ the state space,
\item $A$ the action space,
\item $\gamma$ the discount factor,
\item $R$ the reward function
\end{itemize}

A policy $\pi$ is a map from the states to the actions : $\pi:S\rightarrow A$.\\

The \emph{Reinforcement Learning} problem seeks the policy that maximize the expected discounted reward. When following a policy $\pi$ from state $s\in S$, the expected discounted reward is described by the \emph{value function} $V^\pi$ : \\
\begin{equation}
V^\pi(s) = E\left[\left.\sum\limits_{t=0}^{\infty}\gamma^tr_t\right|s_0=s,\pi\right]
\label{eqn:V}
\end{equation}

The state space $S$ is projected into a feature space generated by the function $\psi : S \rightarrow \mathbb{R}^p$. Both the state space and the action space are projected into a feature space generated by the function $\phi : S\times A \rightarrow \mathbb{R}^k$.\\

The value function is approximated via a matrix of coefficients $\omega$, such as :
\begin{equation}
\hat V^\pi(s) = \arg\max_{a\in A}\omega_\pi^T\phi(s,a)
\end{equation}

The \emph{expert} is an agent whose interactions with the world are recorded as a set of transitions : $D_E = \left\{\{s_i,a_i,s'_i,r_i\}_i\right\}$.\\

$D$ is a random transition source.\\

The feature expectation is a mapping from a policy $\pi$ to a function $\mu_\pi : S \rightarrow \mathbb{R}^p$ which gives for a state $s\in S$ the expected discounted value for the features of the states visited when following policy $\pi$ from state $s$ :
\begin{equation}
\mu_\pi(s) = E\left[\left.\sum\limits_{t=0}^{\infty}\gamma^t\psi(s_t)\right|s_0=s,\pi\right]
\label{eqn:mu}
\end{equation}\\
The feature expectation of the expert is denoted $\mu_E$.\\

The IRL problem seeks the reward the expert is maximizing. The imitation problem seeks the reward which, maximized by one of the many RL tools, will give a behavior as close to the expert's as possible.\\

\citet{abbeel2004apprenticeship}'s algorithm solves the imitation pproblem by approximating this reward with a matrix of coefficients $\theta$ such as :
\begin{equation}
\hat R(s) = \theta^T\phi(s)
\end{equation}

The quality of the imitation is assessed via the error term $||\mu_\pi(s_0)-\mu_E(s_0)||_2$. Given that the norm of $\theta$ is bounded, this error term also gives a feeling of the difference of performance under the estimated reward function $\hat R$ :
\begin{equation}
|V^\pi(s_0)-V^E(s_0)| \leq ||\theta||_2||\mu_\pi(s_0)-\mu_E(s_0)||_2
\label{eqn:vmu}
\end{equation}

The main shortcoming of this algorithm lies in its need for either a model of the environnement or at least a sample source $D$, in order to approximate the feature expectation of both the expert and the current policy via a Monte-Carlo estimator.
\section{LSTD-$\mu$}
\label{sec:lstdmu}
The similarity between equations \ref{eqn:mu} and \ref{eqn:V} (both are Bellman equations) allows the use of the LSTD algorithm in order to estimate the feature expectation of a policy.\\

>>>>

\typeout{Batch, Off-Policy Apprenticeship Learning}
\documentclass{article}
\usepackage{ijcai11}
\usepackage{times}
\usepackage{color}
% the following package is optional:
%\usepackage{latexsym} 
\usepackage{amsthm}
\newtheorem{theorem}{Theorem}
\newtheorem{lemma}{Lemma}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{natbib}
\usepackage{graphicx}
\usepackage{subfigure}
%\usepackage{algorithm2e}
\usepackage{algorithm}
\usepackage{algorithmic}
\newcommand{\argmax}{\operatorname*{argmax}} %\operatorname* pour les op. pouvant admettre des limites...

@@

\section{Background}
A sequential decision problem is often framed as a Markov Decision
Process (MDP)~\citep{puterman1994markov}. A MDP is a tuple
$\{S,A,P,R,\gamma\}$ with $S$ being the state space, $A$ the action
space, $P\in\mathcal{P}(S)^{S\times A}$ the set of Markovian
transition probabilities, $R\in\mathbb{R}^S$ the reward function
(assumed to be absolutely bounded by 1) and $\gamma\in[0,1[$ a
discounting factor. A policy $\pi\in A^S$ maps states to action. The
quality of a policy is quantified by the associated value function
$V^\pi$, which associates to each state the expected and discounted
cumulative reward:
\begin{equation}
  V^\pi(s) = E[\sum_{t=0}^\infty \gamma^t R(s_t)|s_0=s, \pi]
\end{equation}
Dynamic programming aims at finding the optimal policy $\pi^*$, that
is one of the policies associated to the optimal value function,
$V^* = \argmax_\pi V^\pi$, which maximizes the value for each state.
If the model (that is transition probabilities and the reward
function) is unknown, learning the optimal control through
interactions is addressed by reinforcement learning.

For inverse reinforcement learning, the problem is reversed. It is
assume that an expert acts according to an optimal policy $\pi_E$,
this policy being optimal according to some unknown reward function
$R^*$. The goal of inverse reinforcement learning is to learn this
reward function from sampled trajectories of the expert. This is a
difficult and ill-posed problem~\citep{ng2000algorithms}. Apprenticeship
learning through IRL, which is the problem at hand in this paper,
has a somehow weaker objective: it aims at learning a policy
$\hat{\pi}$ which is (approximately) as good as the expert policy
$\pi_E$ under the unknown reward function $R^*$, for a known initial
state $s_0$ (this condition can be weakened by assuming a
distribution for initial states, this is not done here for clarity
of exposition). Now, the approach proposed in~\citep{abbeel2004apprenticeship} is
presented.

We assume that the true reward function belongs to some hypothesis
space $\mathcal{H}_\phi = \{\theta^T \phi(s),
\theta\in\mathbb{R}^p\}$, of which we assume the basis functions to
be bounded by 1: $|\phi_i(s)|\leq 1, \forall s\in S, 1\leq i \leq
p$. Therefore, there exists a parameter vector $\theta^*$ such that:
\begin{equation}
  R^*(s) = (\theta^*)^T \phi(s)
\end{equation}
In order to ensure that rewards are bounded, we impose that
$\|\theta\|_2\leq 1$. For any reward function belonging to
$\mathcal{H}_\phi$ and for any policy $\pi$, the related value
function $V^\pi(s)$ can be expressed as follows:
\begin{align}
  V^\pi(s) &= E[\sum_{t=0}^\infty \gamma^t \theta^T \phi(s_t)|s_0=s, \pi]
  \\
  &= \theta^T  E[\sum_{t=0}^\infty \gamma^t \phi(s_t)|s_0=s, \pi]
\end{align}
Therefore, the value function is also linearly parameterized, with
the same weights and with basis functions being grouped into the
so-called \emph{feature expectation} $\mu^\pi$:
\begin{equation}
  \mu^\pi(s) = E[\sum_{t=0}^\infty \gamma^t \phi(s_t)|s_0=s, \pi]
\end{equation}
Recall that the problem is to find a policy whose performance is
close to that of the expert's for the starting state $s_0$, on the
unknown reward function $R^*$. In order to achieve this goal, it is
proposed in~\citep{abbeel2004apprenticeship} to find a policy $\tilde{\pi}$ such
that $\|\mu^{\pi_E}(s_0)-\mu^{\tilde{\pi}}(s_0)\|_2\leq \epsilon$
for some (small) $\epsilon>0$. Actually, this ensures that the value
of the expert's policy and the value of the estimated policy (for
the starting state $s_0$) are close for \emph{any} reward function
of $\mathcal{H}_\phi$:
\begin{align}
  |V^{\pi_E}(s_0) - V^{\tilde{\pi}}(s_0)| &=
  |\theta^T(\mu^{\pi_E}(s_0)-\mu^{\tilde{\pi}}(s_0))|
  \\
  &\leq \|\mu^{\pi_E}(s_0)-\mu^{\tilde{\pi}}(s_0)\|_2
\end{align}
This last equation uses the Cauchy-Schwartz inequality and the
assumption that $\|\theta\|_2\leq 1$. Therefore, the approach
described here does not ensure to retrieve the true reward function
$R^*$, but to act as well as the expert under this reward function
(and actually under any reward function, notably the unknown optimal
one).

Let us now describe the algorithm proposed in~\citep{abbeel2004apprenticeship} to
achieve this goal:
\begin{enumerate}
  \item Starts with some initial policy $\pi^{(0)}$ and compute
  $\mu^{\pi^{(0)}}(s_0)$. Set $j=1$;
  %
  \item Compute $t^{(j)} = \max_{\theta: \|\theta\|_2\leq 1}
  \min_{k\in\{0,j-1\}}\theta^T(\mu^{\pi_E}(s_0) -
  \mu^{\pi^{(k)}}(s_0))$ and let $\theta^{(j)}$ be the value
  attaining this maximum. At this step, one searches for the reward
  function which maximizes the distance between the value of the
  expert at $s_0$ and the value of \emph{any} policy computed so far
  (still at $s_0$). This optimization problem can be solved using a
  quadratic programming approach or a projection
  algorithm~\citep{abbeel2004apprenticeship};
  %
  \item if $t^{(j)}\leq \epsilon$, terminate. The algorithm outputs a
  set of policies $\{\pi^{(0)}, \dots, \pi^{(j-1)}\}$ among which
  the user chooses manually or automatically the closest to the
  expert (see~\citep{abbeel2004apprenticeship} for details on how to choose this
  policy\footnote{A simple solution, not mentioned in~\citep{abbeel2004apprenticeship},
  is to choose the policy of which the feature expectation is the
  closest to the one of the expert.}).
  Notice that the last policy is not necessary the best;
  %
  \item solve the MDP with the reward function $R^{(j)}(s) =
  (\theta^{(j)})^T\phi(s)$ and denote $\pi^{(j)}$ the associated
  optimal policy. Compute $\mu^{\pi^{(j)}}(s_0)$;
  %
  \item set $j\leftarrow j+1$ and go back to step 2.
\end{enumerate}
There remains three problems: computing the feature expectation of
the expert, solving the MDP and computing feature expectations of
intermediate policies.

As suggested in~\citep{abbeel2004apprenticeship}, solving the MDP can be done
approximately by using any appropriate reinforcement learning
algorithm. In this paper, we use the Least-Squares Policy Iteration
(LSPI) algorithm~\citep{lagoudakis2003least}. There remains to estimate
feature expectations. In~\citep{abbeel2004apprenticeship}, $\mu^{\pi_E}(s_0)$ is
estimated using a Monte Carlo approach over $m$ trajectories:
\begin{equation}
  \hat{\mu}_E(s_0) = \frac{1}{m} \sum_{h=1}^m \sum_{t=0}^\infty
  \gamma^i \phi(s_t^{(h)})
\end{equation}
This approach does not hold if only transitions of the expert are
available, or if trajectories are too long (in this case, it is
still possible to truncate them). For intermediate policies, it is
also suggested to estimate associated feature expectations using a
Monte Carlo approach (if they cannot be computed exactly). This is
more constraining than for the expert, as this assumes that a
simulator of the system is available. In order to address these
problems, we introduce a temporal-difference-based algorithm to
estimate these feature expectations.

\section{LSTD-$\mu$}
\label{sec:lstdmu}
The similarity between equations \ref{eqn:mu} and \ref{eqn:V} (both are Bellman equations) allows the use of the LSTD algorithm in order to estimate the feature expectation of a policy.\\
